---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## crawl-workshop
slides, data and code in support of workshops at AFSC and SWFSC for the R package `crawl`

### Logistics and Schedule

14-16 February, 2017  
9am - 5pm  
MML Conference Room, 2039  
  
NOAA Western Regional Center  
Alaska Fisheries Science Center (Bldg 4)  
7600 Sand Point Way NE, Seattle, WA  

[Identification Requirements for Accessing the WRC](http://www.wrc.noaa.gov/NewIdRequirements.htm)

Nearest Hotel: 
[Silver Cloud (University)](https://www.silvercloud.com/university/)


### Workshop Agenda

#### Day 1: 
1. Introductions/Why are you here?
1. The Theory of Crawling (Devin Johnson)
1. Lunch
1. Practical Guide to Crawling (Josh London)

#### Day 2: 
1. Hands-on w/ example datasets

#### Day 3: 
1. BYODHF (Bring Your Own Data, Have Fun)

### What To Bring and Have Installed

1. A laptop computer with the following software installed
    1. R 3.3.2
    1. RStudio 1.0.136
    1. (Windows users) [RTools34](https://cran.r-project.org/bin/windows/Rtools/)
1. The following R packages from CRAN: `devtools`, `tidyverse`, `lubridate`, `sp`, `raster`, `argosfilter`, `doParallel`, `xts` (see below for installation code)
1. THe following R packages installed from github: `crawl`, `crawlr`, `nPacMaps`
1. AFSC staff should work with IT to insure they are signed up for the EduRoam campus wifi network. Visitors will use the NOAA Guest wifi.
1. A healthy does of curiosity and patience. This will be the first time teaching this workshop, so it won't be perfect

### Installation of R Packages

```{r, r-package-install, eval=FALSE}
library(devtools)

devtools::install_github("NMML/crawl",ref = "devel")
devtools::install_github("jmlondon/nPacMaps")
devtools::install_github("jmlondon/crawlr")
devtools::install_github("jmlondon/wcUtils", ref = "develop")

pkgs <- c(
  "tidyverse",
  "lubridate",
  "sp",
  "raster",
  "doParallel",
  "argosfilter"
)

install.packages(pkgs = pkgs)
```
