---
title: "Theoretical Crawling"
subtitle: "A guide to the mathematics of animal movement"
author: "Devin S. Johnson"
date: "2017/02/05"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(png)
library(grid)
library(crawl)
```


class: middle
# Outline

.pull-left[
## Mathematics

1. Discrete-time random walks
1. Correlated random walks
1. Brownian motion
1. Ornstein-Ulenbeck (OU) process
1. Continuous-time CRW
1. Stochastic differential equations
1. Integrated SDEs
]

.pull-right[
## Statistics

1. Maximum likelihood inference 
1. Bayesian inference
1. State-space models
1. Kalman filter/smoother (KFS)
1. Importance sampling
1. Process imputation
]

---
# For more depth ...

```{r echo=FALSE}
img <- readPNG("Book_Cover_Final.png")
grid.raster(img)
```



---
# Stochastic differential equations

$$\mu(s) = \int_0^t \nu(u) du$$



