---
title: "Practical Crawling"
subtitle: "A guide to fitting models to real data"
author: "Josh M. London"
date: "2017/02/05"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---
class: middle

# Outline

- Assemble Source Data
- Intro to the Tidyverse
- Tidy Data for Telemetry
- Starting Locations
- Activity Data (e.g. haul-out)
- Speed filter

---
class: middle
# Outline

- Model Parameters
- Parallel Process Multiple Animals
- Try-Catch Error Trapping
- Plotting Maps
- Helper Package `crawlr`
- Utilization Distributions
---

# Assemble Source Data

## Start with the rawest form of source data

* Comma/Tab-separated files (Argos CDs)
* Tag Manufacturer Source Data
* Movebank or other Data Repositories

## R has a variety of packages for ingesting all of these data types

* `read.csv()`, `readr::read_csv()`
* `wcUtils` package for Wildlife Computers Data Portal
* `move` package for Movebank.org, other http, API, and rest service functions

---

## Script Everything

### Develop everything within R scripts  
(or R Markdown files)

`load_data.R`,  
`tidy_data.R`,  
`model_data.R`,  
`plot_data.R`

### Key Principles

1. Each script can always be sourced in full
2. Each step in your process is reproducible
3. Create output files that are the input of the next step

---

# Tidy Data for Telemetry

### Tidy Data:

1. Each variable forms a column.
2. Each observation forms a row.
3. Each type of observational unit forms a table.

Philosophy and Collection of Packages from Hadley Wickham

### Additional Resources:

1. https://cran.r-project.org/pub/R/web/packages/tidyr/vignettes/tidy-data.html
2. https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
3. https://www.jstatsoft.org/article/view/v059i10/v59i10.pdf

---
class: middle

## Telemetry Data is Messy

---
class: middle

## Column headers are often values, not variable names.

For example, behavior data from telemetry (e.g. dive bins and histograms).

---
class: middle

## Date-Time Columns are Separated

The Space-Time Continuum Does Not Allow for the Separation of Date & Time

---
class: middle

## Date-Time Formats are Variable

### Make Your Life Easy and Adopt ISO8601

- `2017-02-13 16:02:07+00:00`  
- `2017-02-13 16:02:07Z`  
- `20170213 160207Z`  

### Or, Explicitly Convert Your Date Columns to POSIXct in R

(the `lubridate` package is your friend)

---

## Messy Data Example

### Timeline Percent Dry Data

1. Multiple Hourly Percent-Dry Observations per Row
2. Column Headers Contain Values (e.g. Bin1, Bin2, ...)
3. Non Standard Date-Time Format (`%H:%M:%S %d-%b-%Y`)

---

## Messy Data Example

### Timeline Percent Dry Data

```{r, read-histo-csv, warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)

histo_data <- readr::read_csv('assets/143949-Histos.csv')
histo_data %>% select(1:39) %>% slice(1:8)
```

---

## Tidying Data Example

```{r, tidy-histo-csv, warning=FALSE, message = FALSE}
## Convert Column Names to Lower Case for Easier Coding
colnames(histo_data) <- tolower(colnames(histo_data))

## Specify Our Date Format and Convert to POSIX w/ lubridate
date_format <- "%H:%M:%S %d-%b-%Y"
histo_data <- histo_data %>% 
  dplyr::mutate(date = lubridate::parse_date_time(date,date_format))

## Create a tbl_df that Relates Bin Columns to Day Hours
bins <- tibble(bin=paste("bin",1:24,sep=""),hour=0:23)

## Chain Together Multiple Commands to Create Our Tidy Dataset
tidy_histo <- histo_data %>% 
  dplyr::filter(histtype %in% 
                  c('Percent','1Percent')) %>%
  tidyr::gather(bin,percent_dry, starts_with('bin')) %>%
  dplyr::left_join(bins,by="bin") %>%
  dplyr::mutate(datadatetime = date + lubridate::hours(hour)) %>%
  dplyr::select(deployid,datadatetime,percent_dry) %>%
  dplyr::arrange(deployid,datadatetime)
```

---

## Tidy'd Data Output

```{r, tidy-histo-output}
tidy_histo
```

---

## A Package for Tidying Wildlife Computers Data

### `wcUtils`

https://github.com/jmlondon/wcUtils

Under heavy development, but provides some helper functionality for Wildlife Computers tag data

```{r, tidy-histo, warning=FALSE, message=FALSE, echo = TRUE, eval = FALSE}
library(wcUtils)
histo_data <- wcUtils::read_histos('assets/143949-Histos.csv')
histo_data <- wcUtils::tidyTimelines(histo_data)
```

---
class: middle

## Brief Introduction to `dplyr`

The `dplyr` package is a collection of functions meant to help you manipulate
your data and perform summary functions.

`dplyr` also incorporates use of the `magrittr` package which introduces pipes
in the form of the `%>%` command.

---

## Brief Introduction to `dplyr`

```{r, dplyr-examples, message = FALSE, warning = FALSE}

library(nycflights13)
head(flights)

```

---

## Introduction to `dplyr` Cont'd

### `dplyr` verbs

- `filter()` (and `slice()`)
- `arrange()`
- `select()` (and `rename()`)
- `distinct()`
- `mutate()` (and `transmute()`)
- `summarise()`
- `sample_n()` (and `sample_frac()`)

---
## `filter()`


```{r}
flights %>% filter(., month == 1, day == 1)
```

---

##  `select()`

```{r}
flights %>% select(.,year, month, day)
```

---

## `mutate()`

```{r}
flights %>% mutate(.,
  gain = arr_delay - dep_delay,
  speed = distance / air_time * 60)
```
---
class: middle

## Hands On Live Demonstration

1. PTT 160941 - Harbor Seal from the Aleutian Islands
1. Raw Data File: `examples/data/160941-Locations.csv`
1. Template R Script: `examples/scripts/160941-ReadData.csv`

---
# Starting Locations

- starting values are important to getting a good model fit
- in most cases, release date-time and release coordinates are known
- first step is to merge that information with our imported location data

---

# Activity Data (e.g. haul-out)

---

# Speed Filter

---

# Model Parameters

---

# Try-Catch Error Trapping

---

# Plotting Maps

---

# Helper Package `crawlr`

---

# Utilization Distributions

